---
title: \small Intergenerational Similarity in Musical Preferences
author: Mysterious Group
format:
  dashboard:
    orientation: columns
  
editor: 
  markdown: 
    wrap: 72
---

```{r setup}
#| include: false
library(tidyverse)
df <- read.csv("Data Science for Everyday Music Listening_January 28, 2026_04.48.csv", skip = 2)
```

# Main Page

##  {.sidebar}

Everyday music listening often feels personal and self-directed. People
build playlists that reflect their identity and rarely question where
their musical preferences come from. Yet for many individuals, their
earliest listening experiences occurred before they had any control over
the music, for example while sitting in the backseat of a car as a
parent chose what played. In this project, we address this gap by
examining measurable audio features of music, including energy, valence,
tempo, and danceability. Rather than comparing genres, we analyze
parents’ and children’s favourite songs directly using these features.
Our central research question is: *To what extent do the musical
characteristics of parents’ favourite songs influence those of their
children’ s favourite songs?*

## Column 1 {width="40%"}

## Column 2 {width="60%"}

### Row 1

This is more text. Below it, you'll see another row with a tabset.

### Row 2 {.tabset}

```{r}
#| title: Chart 2
ggplot()
```

```{r}
#| title: Chart 3
ggplot()
```


# Methodology
### Row {.columns}

::: {.column width="33%"}

```{r}
#/ content: valuebox
#/ title: "Total Participants"
list(
  icon = "people-fill",
  color = "primary",
  value = "58"
)
```
:::

::: {.column width="33%"}
```{r}
#/ content: valuebox
#/ title: "Valid Answers"
list(
  icon = "check-circle-fill",
  color = "success",
  value = "48"
)
```
:::

::: {.column width="33%"}
```{r}
#/ content: valuebox
#/ title: "Parent-Child Pairs"
list(
  icon = "house-check-fill",
  color = "info",
  value = "24"
)
```
